---
title: "poli_a2"
author: "XU Qingluo"
date: "2023-10-07"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 0. Load the `tidyverse` Packages

```{r}
library(tidyverse)
```

### 2. Subset by columns

### i. Create a dataset containing only the country-year identifiers and indicators of education quality.

```{r}

d_1789_1827 <- read_csv("_DataPublic_/vdem/1789_1827/vdem_1789_1827_external.csv")

d_1828_1866 <- read_csv("_DataPublic_/vdem/1828_1866/vdem_1828_1866_external.csv")

d_1867_1905 <- read_csv("_DataPublic_/vdem/1867_1905/vdem_1867_1905_external.csv")

d_1906_1944 <- read_csv("_DataPublic_/vdem/1906_1944/vdem_1906_1944_external.csv")

d_1945_1983 <- read_csv("_DataPublic_/vdem/1945_1983/vdem_1945_1983_external.csv")

d_1984_2022 <- read_csv("_DataPublic_/vdem/1984_2022/vdem_1984_2022_external.csv")

d_1789_2022 <- bind_rows(d_1789_1827, d_1828_1866, d_1867_1905, d_1906_1944, d_1945_1983, d_1984_2022)

d <- d_1789_2022 |> select(country_name, year, e_peaveduc, e_peedgini)
```

### ii. Rename the columns of education quality to make them informative.

```{r}

d_renamed <- d |>
  rename("Education" = "e_peaveduc", "Educational_inequality" = "e_peedgini",
         "Country" = "country_name", "Year" = "year") 

d_renamed
```

### 3. **Subset by rows**

### i. List 5 countries-years that have the highest education level among its population.

```{r}

d_highest_education_level <- d_renamed |> distinct(Country, .keep_all = TRUE) |> slice_max(order_by = Education, n = 5) 

d_highest_education_level
```

### ii. List 5 countries-years that suffer from the most severe inequality in education.

```{r}

d_most_severe_inequality <- d_renamed |> distinct(Country, .keep_all = TRUE) |> slice_max(order_by = Educational_inequality, n = 5) 

d_most_severe_inequality
```

### **4.Summarize the data**

### i. Check data availability: For which countries and years are the indicators of education quality available?

```{r}

d_availability <- d_renamed |>
   filter(!is.na(Education), !is.na(Educational_inequality))

d_availability
```

### ii. Create two types of country-level indicators of education quality

### a. Average level of education quality from 1984 to 2022

```{r}

average_education <- d_availability |>
  filter(Year >= 1984 & Year <= 2022) |>
  group_by(Country) |>
  summarise(the_average_education = mean(Education))

average_education
```

```{r}

average_educational_inequality <- d_availability |>
  filter(Year >= 1984 & Year <= 2022) |>
  group_by(Country) |>
  summarise(the_average_educational_inequality = mean(Educational_inequality))

average_educational_inequality
```

### b. Change of education quality from 1984 to 2022

```{r}

change_education <- d_availability |>
  filter(Year >= 1984 & Year <= 2022) |>
  group_by(Country) |>
  summarise(the_change_education = last(Education) - first(Education))

change_education
```

```{r}

change_educational_inequality <- d_availability |>
  filter(Year >= 1984 & Year <= 2022) |>
  group_by(Country) |>
  summarise(the_change_educational_inequality = last(Educational_inequality) - first(Educational_inequality))

change_educational_inequality
```

### iii. Examine the data and briefly discuss: Which countries perform the best and the worst in terms of education quality in the past four decades?

### The best:

```{r}
average_education |>
  slice_max(order_by = the_average_education, n = 1)
```

```{r}
average_educational_inequality |>
  slice_min(order_by = the_average_educational_inequality, n = 1)
```

### The worst:

```{r}
average_education |>
  slice_min(order_by = the_average_education, n = 1)
```

```{r}
average_educational_inequality |>
  slice_max(order_by = the_average_educational_inequality, n = 1)
```
